Automatic discovery of cell types and microcircuitry from neural connectomics

### 2015-02-27 docs/paper.tex
spell check

\documentclass{article}
\usepackage[american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[margin=true,inline=true,index=true, status=draft]{fixme}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{amsmath}
%\usepackage{subcaption}
%\usepackage{multicol}
\usepackage{csquotes}
\usepackage{pdfpages}
\usepackage{placeins}

@@

%\renewcommand{\includegraphics}[2][]{\fbox{}}

@@

\begin{abstract}
  Neural connectomics has begun producing massive amounts of data,
  necessitating new analysis methods to discover the biological and
  computational structure. It has long been assumed that discovering
  neuron types and their relation to microcircuitry is crucial to
  understanding neural function. Here we developed a nonparametric
  Bayesian technique that identifies neuron types and microcircuitry
  patterns in connectomics data. It combines the information
  traditionally used by biologists, including connectivity, cell body
  location and the spatial distribution of synapses, in a
  principled and probabilistically-coherent manner. We show that the
  approach recovers known neuron types in the retina and enables
  predictions of connectivity, better than simpler algorithms. It also
  can reveal interesting structure in the nervous system of
  C. elegans and a historic microprocessor.
  Our approach extracts structural meaning from
  connectomics, enabling new approaches of automatically deriving
  anatomical insights from these emerging datasets.
\end{abstract}

@@

\section*{Introduction}
Emerging connectomics techniques \autocite{Morgan2013,Zador2012}
promise to quantify the location and connectivity of each neuron
within a tissue volume. These massive datasets will far exceed the
capacity of neuroanatomists to manually trace small circuits, thus
necessitating computational, quantitative, and automatic methods for
understanding neural circuit structure.  The impact of this kind of
high-throughput transition has been seen before -- rise of sequencing
techniques necessitated the development of novel computational methods
to understand genomic structure, ushering in  bioinformatics
as an independent discipline \autocite{Koboldt2013}.
\fxwarning{cite background article, cite the paper from Yuste compare supervised unsupervised on morphology, cite seungs new article}

The brain consists of multiple kinds of neurons, each of which is
hypothesized to have a specific role in the overall
computation. Neuron types differ in many ways, e.g. chemical or
morphological, but they also differ in the way they connect to one
another. In fact, the idea of well defined, type-dependent local
connectivity patterns (microcircuits) has been prominent in many
areas, from sensory (e.g. retina, \autocite{Masland2001} to processing
(e.g. neocortex \autocite{Mountcastle1997}) to movement (e.g. spinal
cord) \autocite{Grillner2005}. These sorts of repeated computing
patterns are a common feature of computing systems, even arising in
human-made computing circuits. It remains an important challenge to
develop algorithms to use anatomical data, e.g. connectomics, to
automatically back out underlying microcircuitry.

@@

To address these challenges, here we describe a Bayesian
nonparametric model that can discover circuit structure automatically
from connectomics data: the cell types, their spatial patterns of
interconnection, and the locations of somata and synapses. We show
that by incorporating this additional information, our model both
accurately predicts the connection as well as agrees
with human neuroanatomists as to the identification of cell types.

We primarily focus on the recently-released mouse retina connectome
\autocite{Helmstaedter2013}, but additionally examine the C. elegans
connectome \autocite{White1986}. 
Comparing the cell
types discovered by the algorithms with those obtained manually by
human anatomists reveals a high degree of agreement. We thus present a
scalable probabilistic approach to infer microcircuitry from
connectomics data available today and in the future.

@@

We build a structured probabilistic model which begins with the
generic notion of a cell being a member of a single, unobserved type
-- and these types affect soma depth, distribution of synapses, as
well as a cell type and distance dependent connection probability. For
example, retinal ganglion cells may synapse on nearby, but not far
away, amacrine cells, with ganglion cells being superficial and having
a broad distribution of synapses.

@@

As an illustrative example, consider a network with only three cell
types, labeled A, B, and C. Assume the network has three types of
cells, they are uniformly distributed in space, and that the
probability of connection between any two cells, $c_i$ and $c_j$,
depends only on their type and their distance, according to a logistic
(sigmoidal) function. Let A connect only to nearby B and C cells, but B
connect to any C regardless of distance.

@@

We start with a model for connectivity, the infinite stochastic block
model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
meaningfully cluster connection graphs while learning the number of
hidden groups, or types. We extend this approach by adding distance
dependence to model salient aspects of microcircuitry via logistic and
exponential distance-link functions.  We additionally model cell body
depth unimodally and synapse density profile multimodally (see
Methods for mathematical details).

\fxwarning{Do we talk about cross-validation early on? What about
  posterior uncertainty?}

@@

To assess the accuracy of a clustering compared to that determined by
neuroanatomists, we employ the adjusted Rand index (ARI). ARI is an established metric which measures the similarity of two clusterings -- two identical clusterings will have an ARI of 1.0. There are no perfect clustering comparison metrics as all suffer from the need to balance
``overclustering'' (calling a single type two different types) with
``merging'' (calling two distinct types a single type).  See supplemental for a discussion of
ARI's shortcomings.  Despite its shortcomings we use the ARI as a simple measure of clustering accuracy as it quantitatively captures the difference between the types discovered and the types believed by neuroanatomists.

To assess the accuracy of the model for connections, we use link prediction
accuracy. If our model accurately captures the true structure of
the data, it should be good at predicting if a link exists.
We thus train the model on the data with a subset of the links
marked as unobserved and thus compute our predictive accuracy. We perform 10-way
cross validation on a given dataset, learn the resulting model, and
use that model to predict the missing synapses. Each potential
link between cells is assigned a probability, and we compute
the area under the curve (AUC) for the resulting ROC curve. An AUC of 1.0 means that we perfectly predict the presence and absence of the missing synapses. We use link prediction accuracy to quantify how good the model is at discovering the underlying connectivity.
\fxwarning{Add additional ARI}

@@

\subsection*{Model Mismatch}

\fxwarning{Add xlabel, ylabel, make line width of axes a bit bigger, larger number font, 
all label fonts should be the same size }

To understand how our model would perform on real datasets that do not
perfectly adhere to its generative model we next want to analyze its
performance in such cases. To understand the properties of our model, we
attempt connectivity inference on four sets of synthetic data. This helps us understand what our model would do if the data does not obey our assumptions.

@@

If the data is sampled from our model, inference according to our model, unsurprisingly, is good by all measures. It correctly estimates the number of cell types; it is good at predicting connectivity (high AUC); it agrees with human classification (Rand index); it discovers all types, and leads to homogenous estimates (Fig. \ref{fig:othermodel} first row). If the data comes from a block model without distance dependence, we see that it still does pretty well on all meaningful measures (Fig. \ref{fig:othermodel} second row). It basically estimates correctly that there is no distance dependence. For the mixed membership model (Fig. \ref{fig:othermodel} third row), the model grossly overestimates the number of types, by basically allocating a type for each combination of memberships. Otherwise it still performs relatively well. Lastly, for the latent position clustering model (Fig. \ref{fig:othermodel} fourth row), the model does poorly. If type is continuous instead of discrete then our model is basically trying to cover a continuous set with a discrete scenario leading to rather poor performance. However, as we do expect cell types to be real in biology we might expect our model to do well with real data.

@@

Connectomic efforts so far have reconstructed only small sections of
tissue. Consequently, many connections to cells outside that tissue
volume will be lost. We are concerned that this selective elimination
of connectivity along the boundary might give the appearance of
distance-dependent connectivity when there is none. We thus performed
simulations to check if edge effects could destroy spatial structure
and if edge effects could introduce artificial, spurious spatial
structure. We measure the degree to which distance-dependent effects
can arise from selecting regions that are smaller than the ``scale''
of connectivity (fig ~\ref{fig:edgeeffects}). We do this by generating
two collections of synthetic datasets -- one with distance-dependent
connectivity and one without.  We then in each dataset randomly
examine contiguous circular regions with area varying from zero to the
entire volume, and empirically calculate the spatial variance in
type-dependent connectivity.  We find that, if there is no
distance-dependence, edge effects do not artificially introduce
distance dependence. However, if the section we are examining is too
small, our model can miss the distance dependence. Thus with respect
to distance-dependent connectivity inference, our model errs on the
side of caution. But we also find that for spatial extend that is
similar to the currently available datasets the effects of this are
quite limited.

\fxwarning{Add in a line on this figure to show where the existing data is by looking at old-school tracer reconstruction in the retina. CITE }

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=160mm]{mouseretina2.ai}}
  \caption{Discovering cell classes in the mouse retina connectome. Here
we show the maxium a posteriori (MAP) estimate for the types in the mouse
retina data. 
    a.) Input connectivity data for 950 cells for which soma positions
    were known. b. clustered connectivity matrix, each arbitrary color
    corresponds to a single type, and will be used to identify that type in the remainder
    of the plot. c.)  The spatial distribution of our cell types
    -- each cell type tesselates space. Colors correspond to those in b.)
    d.) Connectivity between our clusters as a function
    of distance -- the cluster consisting primarily of retinal
    ganglion cells (brown node on graph) exhibits the expected near and
    far connectivity.}
\label{fig:mouseretina}
\end{figure}

@@

The algorithm took less than 2 hours to perform inference, dividing
neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
wedge is a type). For each pair of neurons there is a specific
distance dependent connection probability
(fig~\ref{fig:mouseretina}b,c,d), which is well approximated by the
model fit. Moreover, each type of cell is rather isotropically
distributed across space (fig~\ref{fig:mouseretina}e) as should be
expected for true cell types.

@@

As a first measure we compare link prediction accuracy across the
methods (Fig.\ref{fig:mouseretina_compare} B AUC, red).  We find that
given the dataset many techniques allow for good link-predictive
accuracy. All the methods allow decent link prediction with an AUC in
the .9 range. However, our algorithm clearly outperforms the simple
statistical models that only use connectivity.

As a second measure we compare link prediction accuracy across the
methods (Fig.\ref{fig:mouseretina_compare} B ARI, blue). We find that our
algorithm far outperforms the controls. We also find that when it is
based on more of the same information used by anatomists use then it
gets better at agreeing with these anatomists. In particular, using
connectivity, distance, synapse distribution and soma depth leads to
the highest ARI. When using the available information the algorithm
produces a good fit to human anatomist judgments.

Finally we look at the spatial extent of the discovered types both
within a layer and between layers (Fig.\ref{fig:mouseretina_compare}
C). We see that, in the absence of distance information, mere
connectivity information results in types which only span a small
region of space -- essentially “local cliques”. Incorporation of
distance information results in types which span the entire extent of
the layer. The depth variance of all models continues to be
substantially larger than that predicted by human anatomists -- future
directions of work include attempting to more strongly encode this
prior belief of laminarity.

@@

\fxwarning{Fix linewidth too small, align d.e.f. callouts, change multiple fonts. Boxes same size. Need y labels (number of samples in bin) }

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=7.0in]{celegans2.ai}}
  \caption{Discovering connectivity and type in
    \textit{C. elegans}. a.) posterior distribution on cell
    connectivity as a function of discovered type, similar to
    figure~\ref{mouseretina}. In b.) we plot neuroanatomist-derived
    types along with their labels. Our model shows high probability of
    motor neurons, sensory neurons, and various interneuron classes
    being of the same type. Soma positions along the body axis are
    plotted in c.) where we see that we cluster spatially-distributed
    motor neurons together, whereas head sensory neurons are more
    likely to be grouped together as well.  d.) The ROC curves for
    held-out link probability for both the electrical synapses (gap
    junctions) and chemical synapses in c. elegans. e.) The posterior
    distribution of the area under the ROC curve for the curves in
    d. f.) Measurements of the agreement of our identified cell types
    compared to neuroanatomists. The high compleneness but low
    homogeneity (and corresponding low ARI) reflects our model's
    tendency to group multiple types into a single type.}
  \label{fig:celegans}
\end{figure}

@@

\subsection*{Types and connectivity in artificial structures}
\begin{figure}
  \centering 
  \centerline{\includegraphics[width=4in]{mos6502.ai}}
  \caption{Discovering connectivity and type the 6502 microprocessor.
    a.) is the micrograph of the original microprocessor, with the
    region containing the registers under study highlighted. b.) Our
    graph consists of the interconnections of MOS field-effect
    transistors with three terminals, Gate, C1, and C2. Reconstruction
    technique did not permit resolution of C1 and C2 into source and
    drain. c.) The spatial distribution of the transistors in each
    cluster show a clear pattern d.) The clusters and connectivity
    versus distance for connections between Gate and C1, Gate and C2,
    and C1 and C2 terminals on a transistor. Red and teal types have a
    terminal pulled down to ground and mostly function as
    inverters. Purple class are clocked, stateful transistors, orange
    control the ALU and yellow control the special data bus (SDB).}
  \label{fig:mos6502}
\end{figure}

@@

\section*{Measuring clustering similarity}
\fxwarning{Add an explanation of ARI, etc.} 
The adjusted rand index (ARI) is a measure of the similarity of two
different clusterings \autocite{Hubert1985} -- two identical clusters
have an ARI of 1.0 while progressively more dissimilar clusters have
lower ARIs, becoming negative as the clustering gets anti-correlated.

>>>>

\documentclass{article}
\usepackage[american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[margin=true,inline=true,index=true, status=draft]{fixme}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{csquotes}
\usepackage{pdfpages}
\usepackage{placeins}

@@

@@

\begin{abstract}
  Neural connectomics has begun producing massive amounts of data,
  necessitating new analysis methods to discover the biological and
  computational structure. It has long been assumed that discovering
  neuron types and their relation to microcircuitry is crucial to
  understanding neural function. Here we developed a nonparametric
  Bayesian technique that identifies neuron types and microcircuitry
  patterns in connectomics data. It combines the information
  traditionally used by biologists in a principled and probabilistically-coherent manner, including  including connectivity, cell body
  location and the spatial distribution of synapses. We show that the
  approach recovers known neuron types in the retina and enables
  predictions of connectivity, better than simpler algorithms. It also
  can reveal interesting structure in the nervous system of
  C. elegans and a historic microprocessor.
  Our approach extracts structural meaning from
  connectomics, enabling new approaches of automatically deriving
  anatomical insights from these emerging datasets.
\end{abstract}

@@

\section*{Introduction}
Emerging connectomics techniques \autocite{Morgan2013,Zador2012}
promise to quantify the location and connectivity of each neuron
within a tissue volume. These massive datasets will far exceed the
capacity of neuroanatomists to manually trace small circuits, thus
necessitating computational, quantitative, and automatic methods for
understanding neural circuit structure.  The impact of this kind of
high-throughput transition has been seen before -- the rise of sequencing
techniques necessitated the development of novel computational methods
to understand genomic structure, ushering in bioinformatics as an
independent discipline \autocite{Koboldt2013}.

The brain consists of multiple kinds of neurons, each of which is
hypothesized to have a specific role in the overall
computation. Neuron types differ in many ways, e.g. chemical or
morphological, but they also differ in the way they connect to one
another \autocite{Seung2014}. In fact, the idea of well defined,
type-dependent local connectivity patterns (microcircuits) has been
prominent in many areas, from sensory (e.g. retina,
\autocite{Masland2001} to processing (e.g. neocortex
\autocite{Mountcastle1997}) to movement (e.g. spinal cord)
\autocite{Grillner2005}. These sorts of repeated computing patterns
are a common feature of computing systems, even arising in human-made
computing circuits. It remains an important challenge to develop
algorithms to use connectivity-based anatomical data (connectomics), to
automatically back out underlying microcircuitry.  already begun,
focusing on

@@

To address these challenges, here we describe a Bayesian non-parametric
model that can discover circuit structure automatically from
connectomics data: the cell types, their spatial patterns of
interconnection, and the locations of somata and synapses. We show
that by incorporating this additional information, our model both
accurately predicts the connection as well as agrees with human
neuroanatomists as to the identification of cell types. We take as
inspiration previous work on identifying cell types automatically from
morphology \autocite{Guerra2011} and electrophysiology
\autocite{Druckmann2013}.

We primarily focus on the recently-released mouse retina connectome
\autocite{Helmstaedter2013}, but additionally examine the \textit{C. elegans}
connectome \autocite{White1986}.  Comparing the cell types discovered
by the algorithms with those obtained manually by human anatomists
reveals a high degree of agreement. We thus present a scalable
probabilistic approach to infer microcircuitry from connectomics data
available today and in the future.

@@

We build a structured probabilistic model which begins with the
generic notion of a cell being a member of a single type
-- and these types affect soma depth, distribution of synapses, as
well as a cell type and distance dependent connection probability. For
example, retinal ganglion cells may synapse on nearby, but not far
away, amacrine cells, with bipolar cells clearly tesselating space and
synapsing on both.

@@

We start with a model for connectivity, the infinite stochastic block
model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
meaningfully cluster connection graphs while learning the number of
hidden groups, or types. We extend this approach by adding distance
dependence to model salient aspects of microcircuitry via logistic and
exponential distance-link functions.  We form a unimodial model of cell
body depth and a multimodal synapse density profile model (see
Methods for mathematical details).

As an illustrative example, consider a network with only three cell
types, labeled A, B, and C. Assume these cells are uniformly
distributed in space, and that the probability of connection between
any two cells, $c_i$ and $c_j$, depends only on their type and their
distance, according to a logistic (sigmoidal) function. Let A connect
only to nearby B and C cells, but B connect to any C regardless of
distance. This is the prior intuition our model is designed to capture.

For the basic link-distance model, we take as input a connectivity
matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
well as a distance function $d(e_i, e_j)$ representing a (physical)
distance between adjacent cells.  See
the supplemental material for extension to multiple connectivity
matrices. We assume there exist an unknown number $K$ of latent
(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
observed connectivity between two cells $R(e_i, e_j)$ then depends
only on their latent type and their distance through a link function
$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
latent type, $c_i=m$ and $c_j=n$, via a parameter $\eta_{mn}$, as well
as a set of global hyper parameters $\theta$, such that the link
function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.

We then jointly infer the posterior distribution of the
class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
$\eta_{mn}$, and the global model hyperparameters $\theta$ :

@@

In our subsequent analysis makes use of both the full posterior
distribution over these parameters as well as the most probable, or
maximum \textit{a posteriori (MAP), estimate.

For the retina data, we then extend the model with the additional
features indicated. Cell soma depth is modeled as a
cell-type-dependent Gaussian distribution with latent (unknown)
per-type mean and variance. Similarly, each cell has some number $N_i$
of synapses, each of which is drawn from a cell-type-specific density
profile with up to three modes.

Inference is performed via Markov-chain Monte Carlo (MCMC) via three
composable transition kernels -- one for structural, one for per-type
parameters, and one kernel for global parameters and
hyperparameters. Details of data preprocessing, inference parameters,
and runtime can be found in the Supplementary Methods section.

@@

To assess the accuracy of a clustering compared to that determined by
neuroanatomists, we employ three metrics -- clustering homogeneity,
clustering completness, and the adjusted Rand Index. All metrics equal
1.0 when two clusterings completely agree. Homogeneity reflects the
degree to which a found cluster or type contains only a single true
type. Completeness measures how much of a true type is contained
within a single identified type -- a completeness of 1.0 means no true
type is split into multiple subtypes. ARI is a metric that reflects both measures (See
supplemental for more information).

To assess the accuracy of the model for connections, we use link
prediction accuracy. If our model accurately captures the true
structure of the data, it should be good at predicting if a link
exists.  We thus train the model on the data with a subset of the
links marked as unobserved and thus compute our predictive
accuracy. We perform 10-way cross validation on a given dataset \autocite{Guerra2011}, learn
the resulting model, and use that model to predict the missing
synapses. Each potential link between cells is assigned a probability,
and we compute the area under the curve (AUC) for the resulting ROC
curve. An AUC of 1.0 means that we perfectly predict the presence and
absence of the missing synapses. We use link prediction accuracy to
quantify how good the model is at discovering the underlying
connectivity.

@@

\subsection*{Model Mismatch}
We next analyze how our model performs in cases where the data are
generated with assumptions different from ours.  To understand the
properties of our model, we attempt connectivity inference on four
sets of synthetic data. This helps us understand what our model would
do if the data does not obey our assumptions.

@@

If the data is sampled from our model, inference according to our
model, unsurprisingly, is good by all measures. It correctly estimates
the number of cell types; it is good at predicting connectivity (high
AUC); it agrees with human classification (Rand index); it discovers
all types, and leads to homogeneous estimates
(Fig. \ref{fig:othermodel} first row). If the data comes from a block
model without distance dependence, we see that it still does well on
all meaningful measures (Fig. \ref{fig:othermodel} second row). This
is unsurprising, as our model learns the distance dependence, even its
absence. For the mixed membership model (Fig. \ref{fig:othermodel}
third row), the model grossly overestimates the number of types, by
basically allocating a type for each combination of
memberships. Otherwise it still performs relatively well. Lastly, for
the latent position clustering model (Fig. \ref{fig:othermodel} fourth
row), the model does poorly. If type is continuous instead of discrete
then our model is basically trying to cover a continuous set with a
discrete scenario leading to rather poor performance. However, as we
do expect cell types to have a discrete biological basis we might
expect our model to do well with real data.

@@

Connectomic efforts so far have reconstructed only small sections of
neural tissue. Consequently, many connections to cells outside that tissue
volume will be lost. We are concerned that this selective elimination
of connectivity along the boundary might give the appearance of
distance-dependent connectivity when there is none. We thus performed
simulations to check if edge effects could destroy spatial structure
and if edge effects could introduce artificial, spurious spatial
structure. We measure the degree to which distance-dependent effects
can arise from selecting regions that are smaller than the ``scale''
of connectivity (fig ~\ref{fig:edgeeffects}). We do this by generating
two collections of synthetic datasets -- one with distance-dependent
connectivity and one without.  We then in each dataset randomly
examine contiguous circular regions with area varying from zero to the
entire volume, and empirically calculate the spatial variance in
type-dependent connectivity.  We find that, if there is no
distance-dependence, edge effects do not artificially introduce
distance dependence. However, if the section we are examining is too
small, our model can miss the distance dependence. Thus with respect
to distance-dependent connectivity inference, our model errs on the
side of caution. But we also find that for spatial extend that is
similar to the currently available datasets the effects of this are
quite limited.

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=160mm]{mouseretina2.ai}}
  \caption{Discovering cell classes in the mouse retina connectome. Here
we show the maximum a posteriori (MAP) estimate for the types in the mouse
retina data. 
    a.) Input connectivity data for 950 cells for which soma positions
    were known. b. clustered connectivity matrix, each arbitrary color
    corresponds to a single type, and will be used to identify that type in the remainder
    of the plot. c.)  The spatial distribution of our cell types
    -- each cell type tessellates space. Colors correspond to those in b.)
    d.) Connectivity between our clusters as a function
    of distance -- the cluster consisting primarily of retinal
    ganglion cells (brown node on graph) exhibits the expected near and
    far connectivity.}
\label{fig:mouseretina}
\end{figure}

@@

The algorithm took less than 4 hours to perform inference, dividing
neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
wedge is a type). For each pair of neurons there is a specific
distance dependent connection probability
(fig~\ref{fig:mouseretina}b,c,d), which is well approximated by the
model fit. Moreover, each type of cell is rather isotropically
distributed across space (fig~\ref{fig:mouseretina}e) as should be
expected for true cell types.

@@

For our model we show the receiver opererating characteristic curve
(ROC curve) (fig~\ref{fig:mouseretina_compare}d) whcih shows how the
true and false positive rates trade off. We plot the posterior
distribution of the area under this curve in
figure~\ref{fig:mouseretina_compare}e. We then plot the 
posterior distribution for cluster agreement metrics -- completeness,
homogeneity, and ari (figure~\ref{fig:mouseretina_compare}f). We see
that our model tends to over-cluster -- some of the finest-granularity of
neuroanatomist types are grouped as a single type by our model.

We compare link prediction accuracy across the
methods, including our own (Fig.\ref{fig:mouseretina_compare}g AUC, red).  We find that
given the dataset many techniques allow for good link-predictive
accuracy. All the methods allow decent link prediction with an AUC in
the 0.9 range. However, our algorithm clearly outperforms the simple
statistical models that only use connectivity.

As a second measure we compare link prediction accuracy across the
methods (Fig.\ref{fig:mouseretina_compare}g ARI, blue). We find that our
algorithm far outperforms the controls. We also find that when it is
based on more of the same information used by anatomists use then it
gets better at agreeing with these anatomists. In particular, using
connectivity, distance, synapse distribution and soma depth leads to
the highest ARI. When using the available information the algorithm
produces a good fit to human anatomist judgments.

Finally we look at the spatial extent of the discovered types both
within a layer and between layers (Fig.\ref{fig:mouseretina_compare}
h). We see that, in the absence of distance information, mere
connectivity information results in types which only span a small
region of space -- essentially “local cliques”. Incorporation of
distance information results in types which span the entire extent of
the layer. The depth variance of all models continues to be
substantially larger than that predicted by human anatomists -- future
directions of work include attempting to more strongly encode this
prior belief of laminarity.

@@

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=7.0in]{celegans2.ai}}
  \caption{Discovering connectivity and type in
    \textit{C. elegans}. a.) posterior distribution on cell
    connectivity as a function of discovered type, similar to
    figure~\ref{fig:mouseretina_compare}. In b.) we plot
    neuroanatomist-derived types along with their labels. Our model
    shows high probability of motor neurons, sensory neurons, and
    various interneuron classes being of the same type. Soma positions
    along the body axis are plotted in c.) where we see that we
    cluster spatially-distributed motor neurons together, whereas head
    sensory neurons are more likely to be grouped together as well.
    d.) The ROC curves for held-out link probability for both the
    electrical synapses (gap junctions) and chemical synapses in
    c. elegans. e.) The posterior distribution of the area under the
    ROC curve for the curves in d. f.) Measurements of the agreement
    of our identified cell types compared to neuroanatomists. The high
    compleneness but low homogeneity (and corresponding low ARI)
    reflects our model's tendency to group multiple types into a
    single type.}
  \label{fig:celegans}
\end{figure}

@@

\subsection*{Types and connectivity in artificial structures}
\begin{figure}
  \centering 
  \centerline{\includegraphics[width=6in]{mos6502.ai}}
  \caption{Discovering connectivity and type the 6502 microprocessor.
    a.) is the micrograph of the original microprocessor, with the
    region containing the registers under study highlighted. b.) Our
    graph consists of the interconnections of MOS field-effect
    transistors with three terminals, Gate, C1, and C2. Reconstruction
    technique did not permit resolution of C1 and C2 into source and
    drain. c.) The spatial distribution of the transistors in each
    cluster show a clear pattern d.) The clusters and connectivity
    versus distance for connections between Gate and C1, Gate and C2,
    and C1 and C2 terminals on a transistor. Purple and yellow types have a
    terminal pulled down to ground and mostly function as
    inverters. The blue types are clocked, stateful transistors, green
    control the ALU and orange control the special data bus (SDB).}
  \label{fig:mos6502}
\end{figure}

@@

\section*{Measuring clustering similarity}
We compare discovered types to known types via cluster comparison
metrics: cluster homogeneity, cluster completeness, and the adjusted
Rand index. Homogeneity measures how many true types are in a given
found type. If every cell type is split into two types, each subtype
is still completely homogeneous. Completeness measures how many
members of a given true type are split across found types.

The adjusted rand index (ARI) takes into account both effects
\autocite{Hubert1985} -- two identical clusterings have an ARI of 1.0
while progressively more dissimilar clusters have lower ARIs, becoming
slightly negative as the clustering gets anti-correlated.

Figure~\ref{fig:s:comparison} shows the result of taking 20 different clusters
and moving data points between them according to the following operations
\begin{itemize}
\item \textbf{distribute}: take a class and distribute its elements
  uniformly among the remaining types.
\item \textbf{merge} : take a type and merge it into another existing
  type
\item \textbf{split} : take a type and split it into two distinct
  types.
\end{itemize}

We can see the impact on ARI, completeness, and homogeneity as we
perform these operations on more of the original 20 types. In all
cases, ``distribution'' of one type among the others is detremental to
the metric. Splitting impacts completeness but not homogeneity, and
merging impacts homogeneity but not completness.

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=180mm]{cluster_metric_comparison.pdf}}
  \caption{Type agreement evaluation metrics as a function of splitting types, merging types, and randomly distributing cells between types}
  \label{fig:s:comparison}
\end{figure}
